---
title: "Introduction to Leaflet"
author: "Tyler Bonnell"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

pacman::p_load(leaflet,
               leaflet.extras,
               RColorBrewer,
               sp, sf,
               tidyverse,
               ggmap) # ggmap::geocode() to give character address --> output = long/latitude coordinates

```

# Chapter 1: Setting Up Interactive Maps

```{r, Chapter 1}

# Create/load data:
dc_hq <- tibble(hq = c("Datacamp - NYC", "Datacamp - Belgium"),
                lon = c(-73.98575, 4.717863),
                lat = c(40.74856, 50.881363))

# Exercise
dc_hq %>%
leaflet(options = leafletOptions(dragging = TRUE)) %>%
  addProviderTiles("CartoDB") %>%
  addMarkers(lng = dc_hq$lon, lat = dc_hq$lat, popup = dc_hq$hq)


# Setting Default Map View:

## Focus on a single point: 
# setView(lng, lat, zoom)

## Restrict map to a box: 
# fitBounds(lng1, lat1, lng2, lat2) -- lng#/lat# describe two corners of the box which will restrict the map

##  Once user has dragged past a certain point, bounce them back to area of focus: 
# leaflet(options = leafletOptions(dragging = TRUE)) %>% setMaxBounds(lng1, lat1, lng2, lat2)

rm(dc_hq)
```

# Chapter 2: Interactive Web Map of California 4 Year Colleges

IPEDS Dataset:
  - name: Name of each college
  - lng
  - lat
  - state
  - sector_label: Public, Private, or For Profit

Exercise: 
    - Map all Colleges
    - Use circle markers instead of pins
    - Color code markers by ipeds$sector_label

```{r, Chapter 2}

# Create/load data:
ipeds <- read_csv("Data/Leaflet/ipeds.csv") 

# Color Palettes
RColorBrewer::display.brewer.all()

# Creating our own Leaflet Color Palettes
pal <- leaflet::colorFactor(palette = c("red", "blue", "brown"),
                            levels = c("Public", "Private", "For-Profit"))

# leaflet::colorNumeric(palette = "", domain = #, reverse = T/F) # To make numeric/continuous leaflet color palettes

# Exercise
ipeds %>%
  leaflet() %>%
  addProviderTiles("CartoDB") %>%
  addCircleMarkers(radius = 2, # circles instead of pin markers (via addMarkers())
                   lng = ~lng,
                   lat = ~lat,
                   popup = ~paste0("<b>", name, "</b>", 
                                   "<br/>", sector_label),
                   color = ~pal(sector_label),
                   opacity = 0.5) %>% # can use label = instead of popup = if you don't want to click.
addLegend(position = "bottomright", pal = pal, 
          opacity = 0.5, title = "Type of College",
          values = c("Public", "Private", "For-Profit"))


```


# Chapter 3: Leaflet Extras & IPEDS Part II
  - leaflet.extras::addSearchOSM() -- make map searchable
  - leaflet.extras::addReverseSearchOSM() -- geocode w/ mouse click
  - leaflet.extras::addResetMapButton() -- reset map view
  
```{r, Chapter 3}

# Example 
# leaflet() %>%
#   addTiles() %>%
#   leaflet.extras::addSearchOSM() %>% # Adds search bar (data entered: automatically zooms us into the location)
#   leaflet.extras::addReverseSearchOSM() %>% # Gives us lat/lon of searched point
#   addResetMapButton() # Reset Icon below search (quickly return to default view)

# IPEDS Part II
(map <- ipeds %>%
  leaflet() %>%
   addTiles(group = "OSM") %>%
     addProviderTiles("CartoDB", group = "Carto") %>%
     addProviderTiles("Esri", group = "Esri") %>% ## Three base maps for toggling via addLayersControl(basegroups = c())
  setView(lat = 39.8282, lng = -98.5795, zoom =3) %>% # Center of middle of the US
      addLegend(position = "bottomright", pal = pal, 
          opacity = 0.5, title = "Type of College",
          values = c("Public", "Private", "For-Profit")) %>%
   addCircleMarkers(data = ipeds[ipeds$sector_label == "Public", ], ## Three circle markers for toggling via addLayersControl(overlayGroups = c())
                    group = "Public",
                    lat = ~lat,
                    lng = ~lng,
                    radius = 2,
                    opacity = 0.5,
                    color = ~pal(sector_label),
                    popup = ~paste0("<b>", htmltools::htmlEscape(name), "</b>",
                                   "<br/>", sector_label)) %>%
   addCircleMarkers(data = ipeds[ipeds$sector_label == "Private", ],
                    group = "Private",
                    lat = ~lat,
                    lng = ~lng,
                    radius = 2,
                    opacity = 0.5,
                    color = ~pal(sector_label),
                    popup = ~paste0("<b>", htmltools::htmlEscape(name), "</b>",
                                   "<br/>", sector_label)) %>%
   addCircleMarkers(data = ipeds[ipeds$sector_label == "For-Profit", ],
                    group = "For-Profit",
                    lat = ~lat,
                    lng = ~lng,
                    radius = 2,
                    opacity = 0.5,
                    color = ~pal(sector_label),
                    popup = ~paste0("<b>", htmltools::htmlEscape(name), "</b>",
                                   "<br/>", sector_label)) %>%
   addLayersControl(., 
                    overlayGroups = c("Public", "Private", "For-Profit"), ## Overlayed Circle Markers
                    baseGroups = c("OSM", "Carto", "Esri")) %>% ## Overlayed Base Maps
   leaflet.extras::addSearchFeatures(.,
                                     searchFeaturesOptions(zoom = 10),
     targetGroups = c("Public", "Private", "For-Profit")) %>% # Add Search Bar (that searches and zooms in our data specifically)
   leaflet.extras::addResetMapButton()) # Add button to revert back to default zoom
   

```

# Chapter 4: Mapping Polygons
- Looking at which colleges reside in America's most wealthy zip codes

```{r, Chapter 4}
load("Data/Leaflet/wealthiest_zips.Rda")

wealthy_zips <- sf::st_as_sf(wealthy_zips) %>%
  sf::st_transform(., 4326)

wealthy_zips %>%
  leaflet() %>%
  addTiles("OSM", group = "OSM") %>%
  addProviderTiles("CartoDB", group = "Carto") %>%
  addProviderTiles("Esri", group = "Esri") %>%
  addPolygons(weight = 1, fillOpacity = 0.7, 
              color = "green", group = "Wealthy Zipcodes",
              label = ~paste0("Mean Income: ", "$",as.character(prettyNum(mean_income, big.mark =","))),
              highlightOptions = highlightOptions(weight = 5, color = "white", bringToFront = TRUE)) %>%
  addCircleMarkers(data = ipeds[ipeds$sector_label == "Public", ], ## Three circle markers for toggling via addLayersControl(overlayGroups = c())
                    group = "Public",
                    lat = ~lat,
                    lng = ~lng,
                    radius = 2,
                    opacity = 0.5,
                    color = ~pal(sector_label),
                    popup = ~paste0("<b>", htmltools::htmlEscape(name), "</b>",
                                   "<br/>", sector_label)) %>%
   addCircleMarkers(data = ipeds[ipeds$sector_label == "Private", ],
                    group = "Private",
                    lat = ~lat,
                    lng = ~lng,
                    radius = 2,
                    opacity = 0.5,
                    color = ~pal(sector_label),
                    popup = ~paste0("<b>", htmltools::htmlEscape(name), "</b>",
                                   "<br/>", sector_label)) %>%
   addCircleMarkers(data = ipeds[ipeds$sector_label == "For-Profit", ],
                    group = "For-Profit",
                    lat = ~lat,
                    lng = ~lng,
                    radius = 2,
                    opacity = 0.5,
                    color = ~pal(sector_label),
                    popup = ~paste0("<b>", htmltools::htmlEscape(name), "</b>",
                                   "<br/>", sector_label)) %>%
  addLayersControl(baseGroups = c("OSM", "Carto", "Esri"),
                   overlayGroups = c("Public", "Private", "For-Profit", "Wealthy Zipcodes"))

```

  