---
title: "Intermediate Interactive Data Visualization with plotly in R"
author: "Tyler Bonnell"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list = ls())

pacman::p_load(plotly,
               crosstalk,
               tidyverse)

launches <- read_csv("Data/Plotly II/launches.csv")
us_economy <- read_csv("Data/Plotly II/state_economic_data.csv")
```

# Chapter 1: Introduction & Review of Plotly

```{r, Chapter 1}

# Using symbols to reflect groups:

# happy %>%
#   plot_ly(x = ~life.expectancy, y = ~happiness) %>%
#   add_markers(symbol = ~income, symbols = c("circle-open","square-open","star-open","x-thin-open"))

# Bubble Charts (size of marker proportional to population)
us_economy %>%
  filter(year == 2017) %>%
  plot_ly(x= ~gdp, y= ~house_price) %>%
  add_markers(size= ~population,
              color = ~region,
              symbol= ~region,
              marker = list(sizemode = "diameter", 
                            sizeref = 1))  # sizeref > 1 will make all symbols smaller. 

```

# Chapter 2: Animating Graphics

Plotly Animations: Keyframe animations -- a frame is a single interactive plot at one time point. These frames are then combined to create a smooth animation.

- add_markers(frame =., )... frame = argument uses a variable to define each frame and determine how frames will be combined into an animation.
- add_markers(frame=., ids =., )... ids = argument ensures that a particular variable value (say country == "Belgium") will always be represented by the same symbol (across the various frames) in the animation. This is relevant if the format and/or ordering of the data changes between frames (i.e; over time). 

Without setting ids =. , the marker symbols correspond to the row indexes (i.e; the 17th marker is for data in the 17th row) in the underlying data frames. If the format or ordering of the data gets changed between frames -- it could cause sudden jumps in the animation that does not accurately reflect the data (2001: 17th row and marker == "Belgium". In 2002, new countries are added, 2002: 17th row and marker == "Bangladesh". Unfortunately, the 17th marker would jump eratically and not accurately reflect what is happening in Belgium).

## Polishing Animations:
- animation_opts()
- frame = , specifies milliseconds between display of frames
- transition = frame (default), time used to transition between frames. If transition < frame, then the animation will "Pause between frames" aka present static graphics and then transition between them.
- easing = "linear" (default), type of transition used between frames. 
- redraw = TRUE (default), should the graphic be entirely re-drawn upon transition between frames? 

## Animation slider options
- animation_slider()
- hide = TRUE (remove the slider)
- currentvalue = list(prefix=NULL) -- edit slider text: remove slider text prefix.
    - currentvalue = list(font = list(color = "black", size = 40))


```{r, Chapter 2}

# Animated scatterplot
us_economy %>%
  plot_ly(x= ~gdp, y = ~house_price) %>%
  add_markers(size = ~population,
              color = ~region,
              frame = ~year,
              ids= ~state,
              marker = list(sizemode = "diameter"))

# Polishing animations
us_economy %>%
  plot_ly(x= ~gdp, y= ~house_price,
          hoverinfo = "text",
          text = ~paste0(state, " (",year,") ","<br>",
                         "Average House Price: ","<b>","$",prettyNum(round(house_price*1000,0), big.mark=","),"<b>")) %>%
  add_markers(size= ~population,
              color = ~region,
              frame = ~year,
              ids= ~state,
              marker = list(sizemode = "diameter",
                            sizeref = 2)) %>%
  layout(
    xaxis = list(title = "Real GDP (millions USD)", type = "log"),
    yaxis = list(title = "Housing price index")) %>%
  animation_opts(frame = 1250,
                 easing = "bounce",
                 redraw = TRUE) %>%
  animation_slider(currentvalue = list(prefix = NULL,
                                       font = list(color = "black", size= 24)))


## Adding Layers: Add Updating Year Text Layer and Remove Slider
us_economy %>%
  plot_ly(x= ~gdp, y= ~house_price,
          hoverinfo = "text",
          text = ~paste0(state, " (",year,") ","<br>",
                         "Average House Price: ","<b>","$",prettyNum(round(house_price*1000,0), big.mark=","),"<b>")) %>%
  add_text(x = 200000, y = 450, text = ~year, frame = ~year, ids = ~1, #ids = ~1 (to prevent 2004 from flying in very weird...)
           textfont = list(color = toRGB("gray80"), size = 120)) %>% ## ADD TEXT LAYER SHOWING THE YEAR WITHIN THE GRAPH
  add_markers(size= ~population,
              color = ~region,
              frame = ~year,
              ids= ~state,
              marker = list(sizemode = "diameter",
                            sizeref = 2)) %>%
  layout(
    xaxis = list(title = "Real GDP (millions USD)", type = "log"),
    yaxis = list(title = "Housing price index")) %>%
  animation_opts(frame = 1250,
                 easing = "linear",
                 redraw = TRUE) %>%
  animation_slider(hide=TRUE)

```

# Chapter 3: Linking Graphics

```{r, Chapter 3}

## LINKED DOT PLOTS

us2017 <- us_economy %>%
  filter(year == 2017) %>%
  mutate(employ_pcnt = round((employment/(population*1000))*100,1)) %>%
  group_by(state, year) %>%
  slice_head()

# Create a SharedData object from us2017
shared_us <- SharedData$new(us2017)

# Create a scatterplot of house_price vs. home_owners
p1 <- shared_us %>%
  plot_ly(x = ~home_owners, y = ~house_price,
          hoverinfo = "text",
          text = ~paste0("State: ", state, "<br>",
                         "Home Ownership Percent: ",home_owners,"%","<br>",
                         "House Price: ","$",prettyNum(round(house_price*1000), big.mark = ","),"<br>")) %>%
  add_markers(color = ~region) %>%
  layout(xaxis = list(title = "Home Ownership Percent"),
         yaxis = list(title = "House Price (in $1,000)"))
 
# Scatterplot of house_price vs. employment rate
p2 <- shared_us %>%
  plot_ly(x = ~employ_pcnt, y = ~house_price,
          hoverinfo = "text",
          text = ~ paste0("State: ", state,"<br>",
                          "Employment Percent: ",employ_pcnt,"%","<br>",
                          "House Price: ","$",prettyNum(round(house_price*1000), big.mark = ","),"<br>")) %>%
  add_markers(color = ~region) %>%
  layout(xaxis = list(title = "Employment Percent"),
         yaxis = list(title = "House Price (in $1,000)"))
  
# Polish the linked scatterplots
subplot(p1, p2,
        titleX = TRUE, shareY = TRUE) %>% 
  hide_legend() %>%
  highlight(on = "plotly_hover")

# highlight() just for a double click
# highlight(on = "plotly_selected") to highlight anything captured by box or lasso selection




## LINKED DOT PLOT - TIMESERIES

# Create a shared data object keyed by region
shared_region <- crosstalk::SharedData$new(data= us_economy, key= ~ region)
  
# Create a dotplot of avg house_price by region in 2017
dp_chart <- shared_region %>%
  plot_ly(hoverinfo = "text",
          text= ~paste0("Region: ",region,"<br>",
                        "Average Regional House Price: $",prettyNum(round(avg.hpi*1000),big.mark=","))) %>%
   filter(year == 2017) %>%
  group_by(region) %>%
  summarize(avg.hpi = mean(house_price, na.rm = TRUE)) %>%
  add_markers(x = ~avg.hpi*1000, y = ~region, color= ~region) %>%
  layout(yaxis = list(title = ""),
         xaxis = list(title = "Average Regional House Price"))
  
# Code for time series plot
ts_chart <- shared_region %>%
  plot_ly(x = ~year, y = ~house_price*1000,
          hoverinfo = "text",
          text= ~paste0("State: ",state,"<br>",
                        "Region: ",region,"<br>",
                        "House Price: $",prettyNum(round(house_price*1000),big.mark=","))) %>%
  group_by(state) %>%
  add_lines(color= ~region) %>%
  layout(xaxis = list(title = "Year"),
         yaxis = list(title = "House Price"))

subplot(dp_chart, ts_chart,
        titleX = TRUE, titleY= TRUE, margin = 0.05) %>%
  hide_legend() %>%
  highlight(on = "plotly_hover")






```

# Chapter 4: Case Study (Space Launches)

```{r, Chapter 4}

```