---
title: "Intermediate Interactive Data Visualization with plotly in R (Case Study)"
author: "Tyler Bonnell"
date: "`r Sys.Date()`"
output: html_document:
  code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## Install/Load Packages
pacman::p_load(tidyverse,
               plotly, 
               crosstalk,
               dplyr,
               tidyr,
               purrr)

## Load Raw Data
launches <- read_csv("Data/Plotly II/launches.csv")
```

# Exploring the Space Race

### Key Variables

-  launch_year = Year of launch
-  agency = Launching agency
-  state_code = Launching agency's state
-  category = Success (O) or failure (F)
-  agency_type = Type of agency
-  tag = Harvard/COSPAR id of launch
-  JD = Julian date of launch
-  type = Launch vehicle
-  variant = Variant of launch vehicle
-  mission = Mission

```{r launches Key Variables}
glimpse(launches)
```


## (1) Launches Over Time
```{r Launches Over Time}

### table of launches by year
launches_by_year <- launches %>%
  count(launch_year)

# create a line chart of launches over time
launches_by_year %>%
  plot_ly(x = ~launch_year, y = ~n) %>%
  add_lines(fill = "tozeroy") %>% # adds shading below the line
  layout(
    xaxis = list(title = "Year"), 
    yaxis = list(title = "Launches")
  )

# create a bar chart of launches over time
# launches_by_year %>%
#   plot_ly(x = ~launch_year, y = ~n) %>%
#   add_bars() %>%
#   layout(
#     xaxis = list(title = "Year"), 
#     yaxis = list(title = "Launches")
#   )


```

## (2) Space Race Timeline
```{r Space Race Timeline}

# table of launches by year
state_launches <- launches %>% 
  filter(agency_type == "state") %>% 
  count(launch_year, state_code)

# create a ShareData object for plotting
shared_launches <- state_launches %>% 
  SharedData$new(key = ~state_code)

# Create a line chart for launches by state, with highlighting
shared_launches %>%
  plot_ly(x = ~launch_year, y= ~n, color= ~state_code) %>%
  add_lines() %>%
  highlight()

```

## (3) State vs Private Launches
```{r State vs Private Launches}
# table of launches by year and agency type
launches_by_year <- launches %>% count(launch_year, agency_type)

# create a ShareData object for plotting
shared_launches <- launches_by_year %>% 
  SharedData$new(key = ~agency_type)

# create a line chart displaying launches by agency type, with highlighting
shared_launches %>%
  plot_ly(x= ~launch_year, y= ~n, color= ~agency_type) %>%
  add_lines() %>%
  highlight()

```

## (4: Animated) Launches Over Time by State
```{r Animated Launches Over Time, eval = FALSE}
# Complete the state_launches data set
annual_launches <- state_launches %>%
  count(launch_year, state_code) %>%
  complete(state_code, launch_year, fill = list(n = 0))

# Create the cumulative data set
cumulative_launches <- annual_launches %>%
  split(f = .$launch_year) %>%
  accumulate(., ~bind_rows(.x, .y)) %>%
  bind_rows(.id = "frame")

# Create the cumulative animation
cumulative_launches %>%
  plot_ly(x = ~launch_year, y = ~n, color = ~state_code) %>%
  add_lines(frame = ~frame, ids= ~state_code)

```

## (5: Animated) State vs Private Launches
```{r, Animated State vs Private Launches}

## NOT IN DATA CAMP
private_launches <- launches %>% 
  filter(agency_type == "private") %>%
  rename(year = launch_year, agency_name = agency)

# Complete the private_launches data set
annual_launches <- private_launches %>%
  count(year, agency_name) %>%
  complete(agency_name, year, fill = list(n = 0))

# Create the cumulative data set
cumulative_launches <- annual_launches %>%
  split(f = .$year) %>%
  accumulate(., ~bind_rows(.x, .y)) %>%
  bind_rows(.id = "frame")

# Create the cumulative animation
cumulative_launches %>%
  plot_ly(x = ~year, y = ~n, color = ~agency_name) %>%
  add_lines(frame = ~frame, ids = ~agency_name)

```

